Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-15 write-ups/PLDI/paper.tex
Squeeeeeeeeeeeeeeze

@@

\ADComment{This could be shorter.}
However, these are not hard-and-fast rules: a scheduler might employ
heuristics based on historical workloads to avoid thrashing.  For
example, if graphics calls are known to arrive at a certain rate then
the scheduler might decide that it is worth accepting an $\offerkill$
from the cooperative kernel on the assumption that a graphics task
will be requested momentarily.  Similarly, the scheduler might provide
some, but not all, compute units to the cooperative kernel in response
to an $\offerfork$ call if it has reason to believe that the remaining
compute units are likely to be useful for other high priority tasks
that are likely to arrive imminently.  With some additional language extensions, it would be possible for the cooperative kernel to hint to the scheduler if, for example, the kernel is at a point of computation where there is less work to do than usual, so that $\offerkill$ calls could be accepted in order to save energy.

@@

\ADComment{One cooperative, one non-cooperative.  Justification:
  sufficient to evaluate the key use case of a long-running irregular
  algorithm co-existing with a series of short-running tasks.
  Simplifieid task of implementing prototype, which we do not want to
  over-enginneer since it is only a mock-up.}

\ADComment{Restrictions on kernel structure: variables at top of
  kernel are implicitly $\transmit$-annotated, and $\offerkill$,
  $\offerfork$ and $\resizingglobalbarrier$ calls must appear only in
  kernel entry function.  Avoiding $\transmit$ allowed us to reuse the
  clang OpenCL front-end, and restricitng starting points for fork to
  be in main kernel allowed a ``goto''-based approach to redirecting
  control flow; threading this through procedure calls would be
  possible but laborious.  Restricting kill points enables kill to be
  implemented as ``return''; exiting from a procedure call stack would
  require stack unwinding support, akin to throwing an exception.
  Again make point that all these issues are only relevant due to our
  mock-up.}

@@

\ADComment{Explain kernel merge tool briefly.  Explain library
  briefly.}

\TSComment{we should mention that non-cooperative kernels must be
written in a persistent thread style}

>>>>

\newcommand{\kernelmerge}{\textsc{KernelMerge}}

@@

Note that the above guidelines are not hard-and-fast rules.  For
example, a smart scheduler might respond selectively to $\offerkill$
and $\offerfork$ calls by predicting future high-priority workloads
based on past workloads.

@@

For ease of implementation, our current prototype supports
multitasking a single cooperative kernel with a single non-cooperative
kernel (though the non-cooperative kernel can be invoked many times).
Rather than adding language support for $\transmit$ annotations, we
assume that all variables declared at the root lexical scope of a
cooperative kernel are $\transmit$-annotated; this allows our
implementation to piggy-back on top of Clang (see below) without
requiring front-end changes.  Finally, we require that $\offerkill$,
$\offerfork$ and $\resizingglobalbarrier$ are called from the entry
function of a cooperative kernel.  This allows us to use
$\keyword{goto}$ to direct threads joining the cooperative kernel to
the appropriate location, and $\keyword{return}$ to force threads to
leave the computation.  These restrictions still allow us to
experiment with interesting irregular algorithms (see
\mysec\ref{sec:experiments}), and could all be removed with
engineering effort, but this effort might not be well invested since a
full implementation of coopeartive kernels would not use the
megakernel approach we present here.

@@

We have built a tool, \kernelmerge{}, that takes a cooperative kernel
and a non-cooperative kernel and merges them to produce a megakernel
with an embedded GPU scheduler.  We implemented \kernelmerge{} using the Clang LibTooling framework,\footnote{\url{http://clang.llvm.org/docs/LibTooling.html}}
leveraging recent Clang support for OpenCL 2.0~\cite{ClangOpenCL20}.
